---
title: "p8105_hw1_iah2124"
author: "Iris Hart"
date: "2024-09-20"
output: github_document
---

## Homework 1

## Problem 0.2
Loading the penguin data sets
```{r}
data("penguins", package = "palmerpenguins")
```

## Problem 1
Loading tidyverse package - includes dplyr for data manipulation
```{r}
library(tidyverse)
```

Viewing the dataset
```{r}
view(penguins)
```

Size of the dataset
```{r}
num_rows = nrow(penguins)
num_cols = ncol(penguins)
```

Finding the mean flipper length from the penguins data set, excluding any missing values
```{r}
mean(penguins[["flipper_length_mm"]], na.rm = TRUE)
```

This data set contains information about penguins including the species, which island they are from, bill length(mm), bill depth(mm), flipper length(mm), body mass(g), sex, and year. There are 344 observations in the data set, with 8 variables. The mean flipper length is 200.92mm. 

loading ggplot to make a scatter plot
```{r}
library(ggplot2)
```

Creating a scatter plot of flipper length vs bill length
```{r}
ggplot(data = penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point() +
  labs(x = "Bill Length (mm)", y = "Flipper Length (mm)", title = "Scatterplot of Flipper Length vs Bill Length")
```

Saving the scatter plot 
```{r}
ggsave("penguins_scatterplot.png")
```

## Problem 2
Creating a data frame

```{r}
df = data.frame(
  sample_normal = rnorm(10),
  logical_vec = rnorm(10)>0, 
  char_vec = sample(letters, 10, replace = TRUE),
  factor_vec = factor(sample(c("A", "B", "C"), 10, replace = TRUE))
)
```

Viewing the data frame
```{r}
view(df)
```

Means of each variable
```{r}
mean(df %>% pull(sample_normal))
mean(df %>% pull(logical_vec))  
mean(df %>% pull(char_vec))
mean(df %>% pull(factor_vec))
```
You can find the mean of the data frame sample, you can also find the mean of the logical vector. These variables are numeric or logical(true valued at 1 and false values at 0), indicating that a mean value can be calculated. For the variables character vector and factor vector, no mean can be calculated. This is because these variables are neither numeric or logical. 

Explicitly convert variables from one type to another
```{r eval=FALSE, include=TRUE}
mean_logical = mean(df %>% pull(logical_vec) %>% as.numeric(), na.rm = TRUE)
mean_char = mean(df %>% pull(char_vec) %>% as.numeric(), na.rm = TRUE)
mean_factor = mean(df %>% pull(factor_vec) %>% as.numeric(), na.rm = TRUE)
```

You can now take the mean of the logical and factor variables as the converting was able to assign numeric values to the logical and factor variables. However, for the character variable you can not take the mean since no numeric value can be assigned. The output you get for the character vector is NA since it can't be changed to numeric. 


