---
title: "p8105_hw1_iah2124"
author: "Iris Hart"
date: "2024-09-20"
output: html_document
---

## Homework 1

## Problem 0.2

Loading the penguin datasets
```{r}
data("penguins", package = "palmerpenguins")
```

## Problem 1
Loading tidyverse package - includes dplyr for data manipulation
```{r}
library(tidyverse)
```

Size of the dataset
```{r}
num_rows = nrow(penguins)
num_cols = ncol(penguins)
```

Finding the mean flipper length from the penguins data set, excluding any missing values
```{r}
mean_flipper_length = mean(penguins$flipper_length_mm, na.rm = TRUE)
```

This data set contains information about penguins including identifying #, which species island they are from, bill length(mm), bill depth(mm), flipper length(mm), body mass(g), sex, and year. There are 344 observations in the data set, with 8 variables. The mean flipper length is 200.92mm. 

loading ggplot to make a scatter plot
```{r}
library(ggplot2)
```

Creating a scatter plot of flipper length vs bill length
```{r}
ggplot(data = penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point() +
  labs(x = "Bill Length (mm)", y = "Flipper Length (mm)", title = "Scatterplot of Flipper Length vs Bill Length")
```




